---
title: Munich Open Data Oktoberfest Analysis
author: Thomas Schmidt
date: "2018-04-11"
output:
  html_notebook:
    theme: yeti
    highlight: default
  prettydoc::html_pretty:
    theme: github
    highlight: vignette
---

# Introduction

# Aim

* Using the API to import data with R
* First public analysis on my blog
* Learning and getting feedback
* Create some insights about the Oktoberfest


* Mention other analysis

# Methods and Material

[data set side](https://www.opengov-muenchen.de/dataset/oktoberfest/resource/e0f664cf-6dd9-4743-bd2b-81a8b18bd1d2)



# Set up environment

We will start this analysis by loading some required packgages:
```{r}
require(httr, quietly = T)
#require(jsonlite, quietly = T)
require(tidyverse, quietly = T, warn.conflicts = F)
require(gridExtra, quietly = T, warn.conflicts = F)
require(grid, quietly = T, warn.conflicts = F)
```

```{r, echo = F}
# custom colors
fillCol1 <- "#1C86EE"
fillCol2 <- "#b2df8a"
fillCol3 <- "#00CD00"
fillCol4 <- "#a6cee3"

# custom theme
my_theme <- list(
  theme_bw(),
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0))))

knitr::opts_chunk$set(
    fig.width   = 10,      
    fig.height  = 5,      
    fig.align   = 'center', 
    echo        = TRUE)    
```

# Importing data from API

Having the environment set up, we will use the *Munich Open Data API* and the `httr` library to load 
our data from the server and store it as a data.frame within R.   
The resource-id (`e0f664cf-6dd9-4743-bd2b-81a8b18bd1d2`), we will use to get our data via the API,
can be found under additional information (german: "zusätzliche Informationen") at the bottom of the [data set side](https://www.opengov-muenchen.de/dataset/oktoberfest/resource/e0f664cf-6dd9-4743-bd2b-81a8b18bd1d2)

```{r}
url <- "https://www.opengov-muenchen.de"
path <- "api/action/datastore_search?resource_id=e0f664cf-6dd9-4743-bd2b-81a8b18bd1d2"

# get raw results
raw.result <- GET(url = url, path = path)

# convert raw results to list
result <- content(raw.result)

# convert results to table
dt <- bind_rows(result$result$records)
```


Let's take a quick view at our data to see, if the import went well:
```{r}
head(dt)
```


The API data import worked!   

As we see, the columns have german names.
We will change this for all english speaking readers. By doing this, we will also performe some
class- and metric corrections.


```{r}
dt <- dt %>% 
  transmute(id = `_id`,
            year = as.integer(jahr),
            duration = as.integer(dauer),
            beer_cons = as.integer(bier_konsum)*100, # change metric to L
            hendl_cons = as.integer(hendl_konsum), # change metric from 1/2 to 1 chicken
            beer_price = as.numeric(bier_preis),
            hendl_price = as.numeric(hendl_preis),
            visitors_total = as.numeric(besucher_gesamt) * 1000000, # change metric to nr. of people
            visitors_day = as.integer(besucher_tag)*1000) # change metric to nr. of people
```


# Data description

Our data have `r ncol(dt)` columns and `r nrow(dt)` rows. 
The data set contains yearly data from `r min(dt$year, na.rm = T)` to `r max(dt$year, na.rm = T)` on
beer- and hendl (half chicken) consumption. It also provides information about the price of both as 
well as total visitors, mean daily visitors and the duration.

Variable          | Metric
----------------- | ----------
beer consumption  | L
beer price        | €/Liter
hendl consumption | Chicken
hendl price       | €/half chicken
total visitors    | People
daily visitors    | People
  

# Data munging


Now we will start to work with the Oktoberfest data set.
We will add some information like the years when the Bavarian Central Agricultural Festival (ZLF)
took place and other variables, which we will use in our analysis

```{r}
# generate vector with years in which zlf festival took place 
zlf_years <- c(seq(1810,1996, 3), seq(2000, max(dt$year),4))

dt <- dt %>% 
  mutate(zlf = factor(ifelse(year %in% zlf_years, 1, 0)),
         hendl_cons_per_visitor = hendl_cons / visitors_total,
         beer_cons_per_visitor = beer_cons / visitors_total) %>% 
  select(-id) # remove is column
```

# Data analysis

## Beer price and consumption

Every year one of the most discussed topics around the Oktoberfest is the increased beer price.
People are always complaining that the beer is too expensive. Since we have the historical data on
beer prices, our first question on the data set is:

**How did the beer price develope from `r min(dt$year, na.rm = T)` to `r max(dt$year, na.rm = T)`?**

*Note: The currency used in the data is €. All prices before the year 2002 have been transformed
from DM (german currency before 2002) to Euro.*

```{r, echo = F, fig.align = "center"}
dt %>% 
  ggplot() +
  geom_line(aes(x = year, y = beer_price), color = fillCol3, size = 1) +
  geom_point(aes(x = year, y = beer_price), color = fillCol3) +
  geom_vline(xintercept = 2002, lty = 2, alpha = 0.7, size = 0.7) +
  geom_text(x = 2002, y = 5, label = "Euro introduction", angle = 90, vjust = 1.1, cex = 3) +
  scale_y_continuous(breaks = seq(0, max(dt$beer_price), 1),
                     labels = scales::dollar_format(suffix = " €", prefix = "")) +
  scale_x_continuous(breaks = seq(min(dt$year), max(dt$year), 1)) +
  xlab("Year") +
  ylab("Price 1L beer") + 
  ggtitle("Development of beer price") +
  my_theme +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.3))

```

In fact the price for 1L beer increased almost linear over the years. We will try to use a simple
linear regression model to descripe the data:

```{r}
fit <- lm(beer_price ~ year, data = dt)
fit_dm <- lm(beer_price ~ year, data = subset(dt, year < 2002))
fit_euro <- lm(beer_price ~ year, data = subset(dt, year >= 2002))
```

Our model suggests that the beer price increased by 
**`r round(coefficients(fit)[2], 2)` Cents per year**.

Fitting models to the years with different curencies separately, we see a small differens in estimated
price increase per year:  

* In years with DM currency the estimated yearly price increase was about 
**`r round(coefficients(fit_dm)[2], 2)` Cents per year**.
* In the years after the Euro has been introduced, the estimated yearly price increase was 
**`r round(coefficients(fit_euro)[2], 2)` Cents per year**.   



```{r, echo = F, fig.align = "center"}
dt %>% 
  ggplot() +
  geom_line(aes(x = year, y = beer_price), color = fillCol3, size = 1) +
  geom_point(aes(x = year, y = beer_price), color = fillCol3) +
  geom_vline(xintercept = 2002, lty = 2, size = 0.7) +
  geom_text(x = 2002, y = 5, label = "Euro introduction", angle = 90, vjust = 1.1, cex = 3) +
  geom_segment(x = min(dt$year), xend = 2002, y = predict(fit_dm, list(year = min(dt$year))),
               yend = predict(fit_dm, list(year = 2002)), size = 1, alpha = 0.7) +
  geom_text(x = 1987, y = 8,
            label = paste("During the DM currency period\nthe estimated yearly\nprice increase was",
                          round(coefficients(fit_dm)[2],2),
                          "Cent"),
            hjust = 0,
            cex = 3.5) +
  geom_segment(x = 2002, xend = max(dt$year), y = predict(fit_euro, list(year = 2002)),
               yend = predict(fit_euro, list(year = max(dt$year))), size = 1, alpha = 0.7) +
  geom_text(x = 2007, y = 6, 
            label = paste("Since the Euro introduction\nin 2002 the estimated yearly\nprice increase was",
                          round(coefficients(fit_euro)[2], 2),
                          "Cent"),
            hjust = 0,
            cex = 3.5) +
  scale_y_continuous(breaks = seq(0, max(dt$beer_price), 1),
                     labels = scales::dollar_format(suffix = " €", prefix = "")) +
  scale_x_continuous(breaks = seq(min(dt$year), max(dt$year), 1)) +
  xlab("Year") +
  ylab("Price 1L beer") + 
  ggtitle("Development of beer price", "Modeled increase for each currency period") +
  my_theme +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.3))
```


Now that we confirmed that the price increased over the years, the next question which arises is:


**Did this increase have a negative influence on mean beer consumption per visitor?**

In order to answer this question we will start with a visualisation of our data:


```{r, echo = F, fig.align = "center"}
p1 <- dt %>% 
  ggplot() +
  geom_line(aes(x = year, y = beer_price), color = fillCol3, size = 1) +
  geom_point(aes(x = year, y = beer_price), color = fillCol3) +
  scale_y_continuous(labels = scales::dollar_format(suffix = " €", prefix = ""),
                     breaks = seq(0, max(dt$beer_price), 1)) +
  scale_x_continuous(breaks = seq(min(dt$year), max(dt$year), 1)) +
  xlab("Year") +
  ylab("Price 1L beer") + 
  ggtitle("Mean beer consumption per visitor compared to price development") +
  my_theme +  
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        plot.margin = unit(c(5.5,5.5,0,5.5), "points"))

p2 <- dt %>% 
  ggplot() +
  geom_col(aes(x = year, y = beer_cons_per_visitor), fill = fillCol1, width = 0.8, alpha = 0.7,
           color = fillCol1) +
  scale_y_continuous(labels = scales::dollar_format(suffix = " L", prefix = ""),
                     breaks = seq(0, max(dt$beer_cons_per_visitor), 0.25)) +  
  scale_x_continuous(breaks = seq(min(dt$year), max(dt$year), 1)) +
  coord_cartesian(ylim = c(0,max(dt$beer_cons_per_visitor) + 0.1)) +
  xlab("Year") +
  ylab("Beer consumption\nper visitor") +
  my_theme +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.3),
        plot.margin = unit(c(0,5.5,5.5,5.5), "points")) 

grid.newpage()
grid.draw(rbind(ggplotGrob(p1), ggplotGrob(p2), size = "last"))
```

We do not see a decline in mean beer consumption per visitor over the years.
In fact, since around year 1999 it seems that it even has been increasing steadily.

To get an even closer look to the data we will performe a correlation test using the pearson
correlation coefficient:

```{r}
cor.test(dt$beer_cons_per_visitor, dt$beer_price)
```

The test suggest a true correlation of `r round(cor(dt$beer_cons_per_visitor, dt$beer_price),2)`.
the p-value is way below 0.05 and we have a narrow confidence intervall.
If we plot the data, all points seem to be on around one line:

```{r, echo = F, fig.align = "center"}
cor.coef <- round(with(dt, cor(beer_cons_per_visitor, beer_price)), 2)
     
dt %>% 
  ggplot() +
  geom_point(aes(beer_price, beer_cons_per_visitor), color = fillCol4, size = 2) +
  geom_text(x = 5, y = 1.15, label = paste("Correlation:",cor.coef)) +
  scale_y_continuous(labels = scales::dollar_format(suffix = " L", prefix = ""),
                     breaks = seq(0, max(dt$beer_cons_per_visitor), 0.25)) +  
  scale_x_continuous(labels = scales::dollar_format(suffix = " €", prefix = ""),
                     breaks = seq(0, max(dt$beer_price), 1)) +
  ylab("beer consumption\nper visitor") +
  xlab("Price 1L beer") +
  ggtitle("Correlation between beer price and consumption", "Correlation = Causation?") +
  my_theme +
  theme(legend.position = c(0.9,0.9),
        legend.background = element_rect(fill = "white",
                                  size = 0.5, linetype = "solid", 
                                  colour = "black"))
```


This result would suggest that beer consumption increases the higher the beer price gets.
Nevertheless, in my opinion, this is an example that **correlation does not imply causation**.
I think that the increased consumption is a separate phenomena and not caused by increased prices.
Some reasons you can think of might be:   

* Oktoberfest gets more and more popular outside of munich. For people coming from Australia, Sweden,
etc. the beer is still cheap compared to their home and they come to have a great time and a lot of
"Maß" beer.

In conlusion the data set suggests that **the price does not 
seem to have a negative influence on mean beer consumption per visitor**.


## Hendl price and consumption

On the Oktoberfest it is not just about drinking. A lot of people also like to have something 
to eat besides their beer. So what about the hendl consumption and price development?

As before we will first have a look at the price only:


*Notes*:
   
1. *The currency used in the data is €. All prices before the year 2002 have been transformed 
from DM (german currency before 2002) to Euro.*   
2. *Before the year 2000 the data are mean prices at the selling places. Since 2000 the data are only
mean prices inside the tents.*  

```{r, echo = F, fig.align = "center"}
dt %>% 
  ggplot() +
  geom_line(aes(x = year, y = hendl_price), color = fillCol3, size = 1) +
  geom_point(aes(x = year, y = hendl_price), color = fillCol3) +
  scale_y_continuous(labels = scales::dollar_format(suffix = " €", prefix = ""),
                     breaks = seq(0, max(dt$hendl_price), 1)) +
  scale_x_continuous(breaks = seq(min(dt$year), max(dt$year), 2)) +
  xlab("Year") +
  ylab("Price for\n1/2 hendl") +
  ggtitle("Development of hendl price") +
  my_theme +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.3))
```

It seems that there was a huge price increase in year 2000. **But we need to be careful**: As mentioned
in the notes above the data collection method changed in 2000. So we need to compare both periods
separately. In the year 2013 we have some kind of price outlier. Here the price was raised 
a huge amount compared to the steady increase the years before. After that the price droped to a 
level which seems to lie on a line of steady price increase. We will try to model the price 
increase from year 2000 on using a linear regression model.

```{r}
fit <- lm(hendl_price ~ year, data = subset(dt, year >= 2000))
```

The model suggests that the hendl price increased by **`r round(coefficients(fit)[2], 2)` Cents per year** since 2000.

When we add the model to our plot it looks like this:

```{r, echo = F, fig.align = "center"}
dt %>% 
  ggplot() +
  geom_line(aes(x = year, y = hendl_price), color = fillCol3, size = 1) +
  geom_point(aes(x = year, y = hendl_price), color = fillCol3) +
  geom_vline(xintercept = 1999.5, lty = 2) +
  geom_text(x = 2002, y = 6.5,
            label =  paste("Since the year 2000 the estimated\nyearly price increase was",
                          round(coefficients(fit)[2],2),
                          "Cent"),
            hjust = 0) +
  geom_segment(x = 2000, xend = max(dt$year), y = predict(fit, list(year = 2000)),
               yend = predict(fit, list(year = max(dt$year))), size = 1, alpha = 0.7) +
  scale_y_continuous(labels = scales::dollar_format(suffix = " €", prefix = ""),
                     breaks = seq(0, max(dt$hendl_price), 1)) +
  scale_x_continuous(breaks = seq(min(dt$year), max(dt$year), 2)) +
  xlab("Year") +
  ylab("Price for\n1/2 hendl") +
  ggtitle("Development of hendl price", "Modeled increase for years since 2000") +
  my_theme +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.3))
```

Above we saw that the beer price did not have a negative influence on the beer consumption.
We will ask the same question concerning the hendl price and consumption. To get a first overview,
we will plot price development and consumption development in one graphic.

```{r, echo = F, fig.align = "center"}
p1 <- dt %>% 
  ggplot() +
  geom_line(aes(x = year, y = hendl_price), color = fillCol3, size = 1) +
  geom_point(aes(x = year, y = hendl_price), color = fillCol3) +
  geom_vline(xintercept = 1999.5, lty = 2) +
  geom_text(x = 2002, y = 6.5,
            label = "Since 2000 only the prices inside\nthe tents were used for the statistic",
            hjust = 0) +
  scale_y_continuous(labels = scales::dollar_format(suffix = " €", prefix = ""),
                     breaks = seq(0, max(dt$hendl_price), 1)) +
  scale_x_continuous(breaks = seq(min(dt$year), max(dt$year), 2)) +
  xlab("Year") +
  ylab("Price for\n1/2 hendl") +
  ggtitle("Mean hendl consumption compared to price development") +
  my_theme +  
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        plot.margin = unit(c(5.5,5.5,0,5.5), "points"))


p2 <- dt %>% 
  ggplot() +
  geom_col(aes(x = year, y = hendl_cons/visitors_total*100), fill = fillCol1, width = 0.8,
           alpha = 0.7, color = fillCol1) +
  scale_y_continuous(breaks = seq(0, max(dt$hendl_cons/dt$visitors_total*100), 1)) +
  scale_x_continuous(breaks = seq(min(dt$year), max(dt$year), 2)) +
  xlab("Year") +
  ylab("Hendl consumption\nper 100 visitors") +
  my_theme +
  theme(axis.text.x = element_text(angle = 45, vjust = 1.5, hjust = 1.4),
        plot.margin = unit(c(0,5.5,5.5,5.5), "points")) 

grid.newpage()
grid.draw(rbind(ggplotGrob(p1), ggplotGrob(p2), size = "last"))

```


We see that, that the hendl consumption droped in 2001 after slowely decreasing from 1991 on.
After the drop, the consumption stayed at a constant level until a similar drop in 2016 whereas
the price went up. 
For me it looks like the hendl consumption beeing correlated with price until year 2000. 
After that the consumption reached a minimum bound. Maybe at this point the remaining hendl 
consumption is not that price sensitive anymore.

We can check the correlation by performing a correlation test using the Pearson correlation
coefficient:

```{r}
with(subset(dt, year < 2000),cor.test(hendl_cons/visitors_total*100, hendl_price))
with(subset(dt, year >= 2000),cor.test(hendl_cons/visitors_total*100, hendl_price))
```

The correlation coefficients suggest a negatvie correlation for both periods. Nevertheless, 
the tests confirm the null hypothisis which says that there is no true correlation between the 
variables.   
For the period before 2000 there is at least a strong tendency 
(`p = `r with(subset(dt, year < 2000),cor.test(hendl_cons/visitors_total*100, hendl_price))$p.value``)
for a true negative correlation. Despite that, the confidence intervall for the correlation coefficient is very big, which gives us a strong uncertainty.


```{r, echo = F, fig.align = "center"}
cor_before_2000 <- round(with(subset(dt, year < 2000),
                              cor(hendl_cons/visitors_total*100, hendl_price)), 2)
cor_after_2000 <- round(with(subset(dt, year >= 2000),
                             cor(hendl_cons/visitors_total*100, hendl_price)), 2)
     
dt %>% 
  mutate(period = ifelse(year < 2000, "Before 2000", "Since 2000")) %>% 
  ggplot() +
  geom_point(aes(hendl_price, hendl_cons/visitors_total*100, color = period), size = 2) +
  geom_label(x = 10, y = 4.8, label = paste0("Correlation: ",cor_after_2000), fill = fillCol4) +
  geom_label(x = 5, y = 3.8, label = paste0("Correlation: ",cor_before_2000, " *"), fill = fillCol2) +
  scale_color_manual(values = c(fillCol2, fillCol4),
                     guide = guide_legend(title = element_blank())) +
  scale_y_continuous(breaks = seq(0, max(dt$hendl_cons/dt$visitors_total*100), 1)) +
  scale_x_continuous(labels = scales::dollar_format(suffix = " €", prefix = ""),
                     breaks = seq(0, max(dt$hendl_price), 1)) +
  ylab("Hendl consumption\nper 100 visitors") +
  xlab("Price for 1/2 hendl") +
  ggtitle("Correlation between hendl price and consumption") +
  my_theme +
  theme(legend.position = c(0.9,0.85),
        legend.background = element_rect(fill = "white",
                                  size = 0.5, linetype = "solid",
                                  color = "black"))
```

In conclusion we can say that the **price seems to have an influence on the hendl consumption up to
a specific point**. Maybe we could explain the constat level of consumption after reaching that
specific price margin the following:   

Normally people would buy a hendl up to their specific price boundary. That would mean a steady
decrease in consumption with increasing price.   

Nevertheless their are a lot of companies, which rent a table in a tent and invite their employees
to come. Most of the time these companies give a free amount of hendl and beer consumption to their
employes. I think that a big company's price sensitivity concerning hendl and beer is as high as 
the price sensitivity of a normal Oktoberfest visitor. Thus, even with increasing prices the basic 
level of consumption does not change a lot.   


## Bavarian Central Agricultural Festival (ZLF)

In the last part of our analysis we will have a closer look at the ZLF and its influence on visitor
count as well as beer and hendl consumption. 

The ZLF is an agricultural exhibition which takes place in parallel to the Oktoberfest at the
Theresienwiese. Before 1996 the exhibition was held every three years. From there on it has been taken
place every 4 years.


### Did the ZLF bring more visitor to the Oktoberfest?

Since the ZLF is at the same location as the Oktoberfest and a lot of farmers are going there in 
order to see the exhibition, it could have an influense on visitor count. We will futher investigate
this hypothisis by looking at our data:

```{r, fig.align = "center"}
dt %>% 
  ggplot() +
  geom_col(aes(x = year, y = visitors_total, fill = zlf), alpha = 0.7) +
  scale_fill_manual(values = c(fillCol1, fillCol3)) +
  my_theme 

dt %>% 
  ggplot(aes(x = zlf, y = visitors_total)) +
  geom_boxplot(aes(fill = zlf), alpha = 0.6) +
  geom_jitter(position = position_jitter(height = 0, width = 0.2)) +
  scale_fill_manual(values = c(fillCol1, fillCol3)) +
  my_theme 
```

```{r}
# test for normal distribution
shapiro.test(dt[dt$zlf == 1,]$visitors_total)
shapiro.test(dt[dt$zlf == 0,]$visitors_total)
```

```{r}
t.test(visitors_total ~ zlf, data = dt)
```

```{r}
kruskal.test(visitors_total ~ zlf, data = dt)
```


### Did the visitors consume more beer in ZLF years
```{r}
dt %>% 
  ggplot() +
  geom_col(aes(x = year, y = beer_cons/visitors_total, fill = zlf))

dt %>% 
  ggplot(aes(x = zlf, y = beer_cons/visitors_total)) +
  geom_violin() +
  geom_boxplot( alpha = 0.6) +
  geom_jitter(position = position_jitter(height = NULL, width = 0.2))
```


### Did the visitors consume more hendl in ZLF years

```{r}
dt %>% 
  ggplot() +
  geom_col(aes(x = year, y = hendl_cons/visitors_total*100, fill = zlf))

dt %>% 
  ggplot(aes(x = zlf, y = hendl_cons/visitors_total*100)) +
  geom_violin() +
  geom_boxplot( alpha = 0.6) +
  geom_jitter(position = position_jitter(height = NULL, width = 0.2))
```

# Conclusion

* Beer price development 
* BP influence on consumption
* Hendl price development
* Hendl price influene on consumption
* ZLF influence


# Quick plots

```{r}
cor.coef <- round(with(dt, cor(beer_cons/visitors_total, beer_price)), 2)
     
dt %>% 
  ggplot() +
  geom_point(aes(beer_price, beer_cons/visitors_total), color = fillCol4, size = 2) +
  geom_text(x = 5, y = 1.15, label = paste("Correlation:",cor.coef)) +
  scale_y_continuous(labels = scales::dollar_format(suffix = " L", prefix = ""),
                     breaks = seq(0, max(dt$beer_cons/dt$visitors_total), 0.25)) +  
  scale_x_continuous(labels = scales::dollar_format(suffix = " €", prefix = ""),
                     breaks = seq(0, max(dt$beer_price), 1)) +
  ylab("beer consumption\nper visitor") +
  xlab("Price 1L beer") +
  ggtitle("Correlation between beer price and consumption", "Correlation = Causation?") +
  my_theme +
  theme(legend.position = c(0.9,0.9),
        legend.background = element_rect(fill = "white",
                                  size = 0.5, linetype = "solid", 
                                  colour = "black"))
```


```{r}
test <- dt %>% 
  mutate(hendl_cons_change = (hendl_cons/visitors_total - lag(hendl_cons/visitors_total,1)),
         hendl_price_change = (hendl_price - lag(hendl_price, 1))/lag(hendl_price,1)) 

test_fit <- lm(hendl_cons_change ~ hendl_price_change, data = test)
summary(test_fit)

```

```{r}
dt %>% 
  mutate(beer_cons_change = (beer_cons/visitors_total - lag(beer_cons/visitors_total,1)),
         beer_price_change = (beer_price - lag(beer_price, 1))/lag(beer_price,1)) %>% 
  ggplot() +
  geom_point(aes(x = beer_price_change, y = beer_cons_change))
```


## beer price increase before and after euro
```{r}
dt %>% 
  ggplot(aes(x = currency, y = beer_price_change/beer_price)) +
  geom_violin() +
  geom_boxplot(alpha = 0.6) +
  geom_jitter()
```



# Ideas

* Zusätzlicher Datensatz mit Zeltpreisen suchen
* Zusätzlicher Datensatz mit Wetter -> % der Tage mit Regen

* Danksagung and munich open data (auch für Zusatzinformationen)


